---
title: "Visualizing the Forest Conversion"
subtitle: "INFO 526 - Project 1"
author: 
  - name: "The Plotting Pandas - Megan, Shakir, Maria, Eshaan, Bharath"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Uncovering Global Deforestation and Soy Bean Consumption - need to change"
format:
   html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
  echo: false
---

## Installing necessary libraries

```{r load_packages, message=FALSE, include=FALSE}
# GETTING THE LIBRARIES
if (!require(pacman))
  install.packages(pacman)


pacman::p_load(tidyverse,
               gridExtra,
               tidytuesdayR,
               dplyr,
               janitor,
               dlookr,        # Exploratory data analysis
               here,          # Standardizes paths to data
               formattable,
               ggpubr,
               maps,
               plotly,
               gganimate,
               imager,
               magick)
```

## Setup

```{r ggplot_setup, message=FALSE, include=FALSE}
#| code-summary: "ggplot theme setup"
# setting theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14, base_family = "sans"))

# setting width of code output
options(width = 65)

# setting figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8,        # 8" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 1,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 150             # higher dpi, sharper image
)
```

## Pulling data from the `tidytuesday` github repository

```{r load_dataset, message=FALSE, include=FALSE}
# Getting the Data using the tidytuesdayR package 
deforestation_data <- tidytuesdayR::tt_load(2021, week = 15)

# Getting all the underlying data in the dataset
forest        <- deforestation_data$forest
forest_area   <- deforestation_data$forest_area
brazil_loss   <- deforestation_data$brazil_loss
soybean_use   <- deforestation_data$soybean_use
vegetable_oil <- deforestation_data$vegetable_oil
total_forest <- read.csv ("https://raw.githubusercontent.com/INFO526-DataViz/project-01-The-Plotting-Pandas/main/data/forest-area-km.csv")

# Data is read to deforestation_by_source from a raw csv file which is in github , as it is not being downloaded from the tidytuesdayR package.
deforestation_by_source <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/deforestation_by_source.csv')
```

## Abstract

\<Abstract\>

## Introduction

\<Introduction\>

## Question 1

### Introduction

The following plots show the change in forest cover from `1980 - 2015` using a map plot. We can observe a significant increase in forest cover `China`, `India`, `Russia`, `USA`, and `Australia` over the past 20 years whereas `Brazil`, `Tanzania` have lost more forest covered area over the past 35 years.

### Approach

\<Approach\> Fetching the world map using `maps` library

```{r world_data, message=FALSE, include=FALSE}
# World data from maps
world <- map_data("world")

# Extracting unique countries from the world dataset and storing them in a tibble
unique_countries <- world |>
  select(region) |>       # region is the column which is having country names info
  unique() |>             # getting unique country names
  as_tibble()

# Removing Antarctica from the world data
world <- subset(world, region != "Antarctica")
```

Cleaning the `forest.csv` and processing data for plots

```{r process_forest_function, message=FALSE, include=FALSE}
# function to pre process the forest dataset
# input  : dataset - tibble
#          unique_countries - tibble
# output : filtered_data - tibble

process_forest <- function (dataset, unique_countries) {
  filtered_data <- dataset |>
    # filtering only entity, year and net_forest_conversion columns
    select(entity, year, net_forest_conversion) |>
    # getting all the countires which are not present in forest dataset for a specific years
    # bind_rows() is used combine combine rows of two data frames
    bind_rows(
      # anti_join() is used to return only the rows from the first dataset that isn't having matching rows in the second dataset based on specified key columns
      anti_join(unique_countries, dataset, by = c("region" = "entity")) |>
        # adding year and net_forest_conversion for that specific year as NA
        mutate(year = dataset[1, "year"]$year, net_forest_conversion = NA)
    ) |>
    # renaming USA and UK so that both these countries are matching in world dataset and forest dataset
    mutate(
      entity = case_when(
        entity == "United States" ~ "USA",
        entity == "United Kingdom" ~ "UK",
        TRUE ~ entity
      )
    ) |>
    # creating a categorical variable forest_converstion to group countries based on their forest conversion
    mutate(
      entity = coalesce(entity, region),
      forest_converstion = case_when(
        net_forest_conversion < -400000 ~ "<-400k",
        net_forest_conversion < -200000 ~ "-400k to -200k",
        net_forest_conversion < -100000 ~ "-200k to -100k",
        net_forest_conversion < 0 ~ "-100k to 0",
        net_forest_conversion < 100000 ~ "0 to 100k",
        net_forest_conversion < 200000 ~ "100k to 200k",
        net_forest_conversion < 400000 ~ "200k to 400k",
        is.na(net_forest_conversion) ~ NA_character_,
        TRUE ~ ">400k"
      )
    ) |>
    # ordering forest_converstion column using factors based on the created categories
    mutate(
      forest_converstion = as_factor(forest_converstion) |>
        fct_relevel(
          "<-400k",
          "-400k to -200k",
          "-200k to -100k",
          "-100k to 0",
          "0 to 100k",
          "100k to 200k",
          "200k to 400k",
          ">400k"
        )
    )
  return(filtered_data)
}
```

Combining data in a decade format

```{r split_forest_data, message=FALSE, include=FALSE}
# Dividing data into tibbles based on year and it creates a list of tibbles
forest_decades <- split(forest, f = forest$year)

# Implementing the pre processing function on top of the created split tibbles
# lapply() is used for applying a function on top of any list
filtered_forests <- lapply(forest_decades, function(forest_ds) {
  process_forest(forest_ds, unique_countries)
})
```

## Filtering countries with a Major change over past `four` decades

```{r highlightingCountriesWithmajorChange, message=FALSE, include=FALSE}
# function to filter countries which are having noteworthy forest conversion
# input  : dataset - tibble
#          variable - column character
# output : highlight_data - tibble

filter_countries <- function (dataset, variable){
  highlight_data <- subset(dataset, variable %in% c("Brazil", "Tanzania", "China", "India", "Russia", "USA", "Australia"))
  return(highlight_data)
}

# filtering countries from world dataset using filter_countries
highlight_world <- filter_countries(world, world$region)

# filtering countries from filtered_forests dataset using filter_countries
highlight_filtered_data <- lapply(filtered_forests, function(forest_ds) {
  filter_countries(forest_ds, forest_ds$entity)
})
```

Mapping colours based on the change in `forest cover area`

```{r colourMapping, message=FALSE, include=FALSE}
# color mapping for different forest conversion categories
color_mapping <-
  c(
    "<-400k"         = "#d73027",
    "-400k to -200k" = "#f46d43",
    "-200k to -100k" = "#fdae61",
    "-100k to 0"     = "#fee08b",
    "0 to 100k"      = "#d9ef8b",
    "100k to 200k"   = "#a6d96a",
    "200k to 400k"   = "#66bd63",
    ">400k"          = "#1a9850"
  )
```

### Analysis

\<Analysis\> Plotting change in forest cover for the decade `1990`

```{r plotFor1990, message=FALSE}
# Creating the ggplot map plot
# Using the filtered_forests$`1990` dataset created earlier as a data source
# using entity as map_id for first layer
# using forest_convestion as fill aesthetic and word as map for second layer
# using highlight_filtered_data$`1990` as another dataset for creating another map layer
# using entity as map_id,forest_convestion as fill aesthetic and highlight_world as map for third layer

plot_1990 <- ggplot(filtered_forests$`1990`, aes(map_id = entity)) +
  geom_map(
    aes(fill  = forest_converstion),
    map       = world,
    color     = "#B2BEB5",
    linewidth = 0.25,
    linetype  = "blank"
  ) +
  geom_map(data = highlight_filtered_data$`1990`, 
           aes(map_id = entity, 
               fill   = forest_converstion), 
           map   = highlight_world, 
           color = "#71797E", 
           show.legend = F) +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_manual(values = color_mapping, na.value = "#F2F3F4") +
  coord_fixed(ratio = 1) +
  labs(title    = "Net Forest Conversion by Country in 1990",
       subtitle = "Net change in forest area measures forest expansion minus deforestation",
       caption  = "Data source: Our World in Data",
       fill     = "Net Forest Conversion (hectares)"
       ) +
  theme_void() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    plot.title       = element_text(size = 19, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 15, color = "azure4", hjust = 0.5),
    plot.caption     = element_text(size = 12, color = "azure4", hjust = 0.95)
  ) +
  guides(
    fill = guide_legend(
      nrow = 1,
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      label.position = "bottom",
      label.hjust = 1,
      label.vjust = 1,
      label.theme = element_text(lineheight = 0.25, size = 9),
      keywidth = 1,
      keyheight = 0.5
    ),
    color = guide_legend("No data", override.aes=list(fill="#F2F3F4"))
  )

ggsave(plot_1990, 
       filename = "images/q1/forest_plot_1990.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

## Plotting change in forest cover for the decade `2000`

```{r plotFor2000, message=FALSE}
# Creating the ggplot map plot
# Using the filtered_forests$`2000` dataset created earlier as a data source
# using entity as map_id for first layer
# using forest_convestion as fill aesthetic and word as map for second layer
# using highlight_filtered_data$`2000` as another dataset for creating another map layer
# using entity as map_id,forest_convestion as fill aesthetic and highlight_world as map for third layer

plot_2000 <- ggplot(filtered_forests$`2000`, aes(map_id = entity)) +
  geom_map(
    aes(fill  = forest_converstion),
    map       = world,
    color     = "#B2BEB5",
    linewidth = 0.25,
    linetype  = "blank"
  ) +
  geom_map(data = highlight_filtered_data$`2000`, 
           aes(map_id = entity, 
               fill   = forest_converstion), 
           map   = highlight_world, 
           color = "#71797E", 
           show.legend = F) +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_manual(values = color_mapping, na.value = "#F2F3F4") +
  coord_fixed(ratio = 1) +
  labs(title = "Net Forest Conversion by Country in 2000",
       subtitle = "Net change in forest area measures forest expansion minus deforestation",
       caption = "Data source: Our World in Data",
       fill = "Net Forest Conversion (hectares)"
       ) +
  theme_void() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    plot.title       = element_text(size = 19, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 15, color = "azure4", hjust = 0.5),
    plot.caption     = element_text(size = 12, color = "azure4", hjust = 0.95)
  ) +
  guides(
    fill = guide_legend(
      nrow = 1,
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      label.position = "bottom",
      label.hjust = 1,
      label.vjust = 1,
      label.theme = element_text(lineheight = 0.25, size = 9),
      keywidth = 1,
      keyheight = 0.5,
    )
  )

plot_2000
ggsave(plot_2000, 
       filename = "images/q1/forest_plot_2000.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

## Plotting change in forest cover for the decade `2010`

```{r plotFor2010, message=FALSE}
# Creating the ggplot map plot
# Using the filtered_forests$`2010` dataset created earlier as a data source
# using entity as map_id for first layer
# using forest_convestion as fill aesthetic and word as map for second layer
# using highlight_filtered_data$`2010` as another dataset for creating another map layer
# using entity as map_id,forest_convestion as fill aesthetic and highlight_world as map for third layer

plot_2010 <- ggplot(filtered_forests$`2010`, aes(map_id = entity)) +
  geom_map(
    aes(fill  = forest_converstion),
    map       = world,
    color     = "#B2BEB5",
    linewidth = 0.25,
    linetype  = "blank"
  ) +
  geom_map(data = highlight_filtered_data$`2010`, 
           aes(map_id = entity, 
               fill   = forest_converstion), 
           map   = highlight_world, 
           color = "#71797E", 
           show.legend = F) +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_manual(values = color_mapping, na.value = "#F2F3F4") +
  coord_fixed(ratio = 1) +
  labs(title = "Net Forest Conversion by Country in 2010",
       subtitle = "Net change in forest area measures forest expansion minus deforestation",
       caption = "Data source: Our World in Data",
       fill = "Net Forest Conversion (hectares)"
       ) +
  theme_void() +
  theme(
    legend.position   = "bottom",
    legend.direction  = "horizontal",
    plot.title       = element_text(size = 19, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 15, color = "azure4", hjust = 0.5),
    plot.caption     = element_text(size = 12, color = "azure4", hjust = 0.95)
  ) +
  guides(
    fill = guide_legend(
      nrow = 1,
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      label.position = "bottom",
      label.hjust = 1,
      label.vjust = 1,
      label.theme = element_text(lineheight = 0.25, size = 9),
      keywidth = 1,
      keyheight = 0.5,
    )
  )

plot_2010
ggsave(plot_2010, 
       filename = "images/q1/forest_plot_2010.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

## Plotting change in forest cover for the decade `2015`

```{r plotFor2015, message=FALSE}
# Creating the ggplot map plot
# Using the filtered_forests$`2015` dataset created earlier as a data source
# using entity as map_id for first layer
# using forest_convestion as fill aesthetic and word as map for second layer
# using highlight_filtered_data$`2015` as another dataset for creating another map layer
# using entity as map_id,forest_convestion as fill aesthetic and highlight_world as map for third layer

plot_2015 <- ggplot(filtered_forests$`2015`, aes(map_id = entity)) +
  geom_map(
    aes(fill  = forest_converstion),
    map       = world,
    color     = "#B2BEB5",
    linewidth = 0.25,
    linetype  = "blank"
  ) +
  geom_map(data = highlight_filtered_data$`2015`, 
           aes(map_id = entity, 
               fill   = forest_converstion), 
           map   = highlight_world, 
           color = "#71797E", 
           show.legend = F) +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_manual(values = color_mapping, na.value = "#F2F3F4") +
  coord_fixed(ratio = 1) +
  labs(title = "Net Forest Conversion by Country in 2015",
       subtitle = "Net change in forest area measures forest expansion minus deforestation",
       caption = "Data source: Our World in Data",
       fill = "Net Forest Conversion (hectares)"
       ) +
  theme_void() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    plot.title       = element_text(size = 19, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 15, color = "azure4", hjust = 0.5),
    plot.caption     = element_text(size = 12, color = "azure4", hjust = 0.95)
  ) +
  guides(
    fill = guide_legend(
      nrow = 1,
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      label.position = "bottom",
      label.hjust = 1,
      label.vjust = 1,
      label.theme = element_text(lineheight = 0.25, size = 9),
      keywidth = 1,
      keyheight = 0.5,
    )
  )

ggsave(plot_2015, 
       filename = "images/q1/forest_plot_2015.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

## Saving our Map Plots as GIFs using `image_animate`

```{r savingas_GIF, message=FALSE}
# making gif using gganimate pacakge

forest_plots <- list.files(path = "images/q1/", full.names = TRUE)
forest_plot_list <- lapply(forest_plots, image_read)

# joining all the saved images
joined_plots <- image_join(forest_plot_list)

# animating the images using image_animate() and restting the resolution
# setting fps = 1
forest_animation <- image_animate(image_scale(joined_plots, "3000x3000"), fps = 1)

# saving image to git
image_write(image = forest_animation, path = "images/world_forestaion.gif")
forest_animation
```

### Discussion

\<Discussion\>

## Question 2

### Introduction

### Approach

\<Approach\> Cleaning and processing of the dataset includes the following steps:

Soybean production in Brazil:

1.  Created a new column for calculating the total soybean production. Removing totals of countries whose total production is 0 using the `subset` function.
2.  Filtering for Brazil under the `entity` column, and `year` between 1990 and 2013.

Forest coverage in Brazil:

1.  Filtering for `year` between 1990 and 2013 in the `forest_brazil` dataset.
2.  There is a parameter 'World' which shows the overall forest coverage data. Filtering out `entity` as 'World' and `year` between 1990 and 2013, and grouping by year.
3.  Using `left_join` we merge the two tables based on the `year` column.
4.  Since we now have percentage data and total data per year, we can calculate the change in forest coverage for Brazil by doing `forest_area.x` \* `forest_area.y`.

```{r q2_brazil_process,warning= FALSE, message=FALSE}

#Cleaning total_forest table
total_forest_cleaned <- clean_names(total_forest)

#Making a new column to calculate the total soybean production
soybean <- soybean_use |>
  mutate(total = human_food + animal_feed + processed)
#Some countries do not have production, and shows as 0. 
#Removing the rows if total=0
soybean <- subset(soybean, total != 0)

# Filter data for Brazil
soybean_brazil <- soybean |>
  filter(entity == "Brazil", year>= 1990&year<=2013)

# Filter data for Brazil forest: 
forest_brazil <- forest_area |>
  filter(entity == "Brazil",year>=1990&year<=2013)

#Finding total forest coverage per year
total_forest_world <- total_forest_cleaned |>
  filter(year >= 1990, year <= 2013, entity == "World") |>
  group_by(year)

# Left join to add total world forest coverage to the forest_brazil dataset
forest_brazil <- forest_brazil |>
  left_join(total_forest_world, by = "year")
 
#Finding actual total coverage for Brazil (percentage * total)
forest_brazil <- forest_brazil|>
  mutate(forest_area_brazil = forest_area.x * forest_area.y / 100)

```

### Analysis

\<Analysis\> Plotting the soybean data

```{r q2_plot_soybean,warning= FALSE, message=FALSE}

# Create a line plot for Brazil soybean production
plot_soybean_brazil <- ggplot(soybean_brazil, aes(x = year, y = total, color = "Brazil")) +
  geom_line(size = 2) +
  geom_point(color = "black") +
  labs(x = "Year", y = "Total (in lb)", title = "Total Soybean Usage Trend in Brazil", caption = "Jon Harmon | TidyTuesday") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::label_number(scale = 1e-06, suffix = "M")) +
  scale_color_manual(values = c("Brazil" = "cornsilk3")) +
  scale_x_continuous(limits = c(1990, 2013), breaks = seq(1990, 2013, by = 2)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
plot_soybean_brazil

ggsave(plot_soybean_brazil, 
       filename = "images/q2/plot_soybean_brazil.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

Animating the soybean production graph to show the trends:

```{r q2_plot_soybean_anim,warning= FALSE, message=FALSE}

# Animate the plot
anim_plot_soybean <- plot_soybean_brazil + transition_reveal(year)

# Save as an animated GIF
anim_save("soybean_brazil_animation.gif", anim_plot_soybean, renderer = gifski_renderer())

# Load the animated GIF
brazil_soybean_animation <- image_read("soybean_brazil_animation.gif")

# Display the animation
brazil_soybean_animation |>
  image_animate(fps = 25) |>
  image_scale("600x400") # Adjust the dimensions as needed

# saving image to git
image_write(image = brazil_soybean_animation, path = "images/soybean_brazil_animation.gif")
```

Plotting the forest coverage in Brazil

```{r q2_plot_forest,warning= FALSE, message=FALSE}

# Create a line plot for Brazil with points
plot_forest_brazil <- ggplot(forest_brazil, aes(x = year, y = forest_area_brazil, color = "Brazil")) +
  geom_line(size = 2) +
  geom_point(color="blue") + 
  labs(x = "Year", y = "Forest coverage (in hectares)", title = "Total forest cover trend in Brazil", caption= "Jon Harmon | TidyTuesday") +
  theme_minimal() +
  theme(legend.position = "none")+
  scale_color_manual(values = c("Brazil" = "red3"))+
  scale_x_continuous(limits = c(1990, 2013), breaks = seq(1990, 2013, by = 2))+
  scale_y_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

plot_forest_brazil

ggsave(plot_forest_brazil, 
       filename = "images/q2/forest_plot_brazil.jpg", 
       height   = 8, 
       width    = 15, 
       unit     = "in", 
       dpi      = 120)
```

```{r q2_plot_forest_anim,warning= FALSE, message=FALSE}

# Animate the plot
anim_plot_forest <- plot_forest_brazil + transition_reveal(year)

# Save as an animated GIF
anim_save("forest_brazil_animation.gif", anim_plot_forest, renderer = gifski_renderer())

# Load the animated GIF
brazil_forest_animation <- image_read("forest_brazil_animation.gif")

# Display the animation
brazil_forest_animation |>
  image_animate(fps = 25) |>
  image_scale("600x400") # Adjust the dimensions as needed

# saving image to gift
image_write(image = brazil_forest_animation, path = "images/forest_brazil_animation.gif")
```

### Discussion

\<Discussion\>

## References
