---
title: "Visualizing the Impact of Deforestation"
subtitle: "Proposal"
format: html
editor: visual
code-annotations: hover
categories: 
  - Proposal
  - DataViz
  - TidyTuesday
---

## Packages Setup

```{r load_pkgs, message=FALSE, warning=FALSE, output=FALSE}
### GETTING THE LIBRARIES
if (!require(pacman)) #<1>
  install.packages(pacman)


pacman::p_load(tidyverse,
               gridExtra,
               tidytuesdayR,
               dplyr,
               dlookr,        # Exploratory data analysis
               DMwR2, 
               here,          # Standardizes paths to data
               formattable,
               ggpubr)   #<2>
```

1.  This block is to check if `pacman` package is installed or not, if it isn't then the package will be installed.

2.  `p_load` is a function from `pacman` package to load other packages, if isn't installed `pacman` will handle the installation.

## Dataset

```{r load_dataset, message=FALSE, warning=FALSE}
# Getting the Data using the tidytuesdayR package 
deforestation_data <- tidytuesdayR::tt_load(2021, week = 15)

# Getting all the underlying data in the dataset
forest        <- deforestation_data$forest
forest_area   <- deforestation_data$forest_area
brazil_loss   <- deforestation_data$brazil_loss
soyabean_use  <- deforestation_data$soybean_use
vegetable_oil <- deforestation_data$vegetable_oil
deforestation_by_source <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/deforestation_by_source.csv')
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

### Forest Data

```{r forest_basic_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
forest |>
  head() |>
  formattable()             #<1>

# The properties of the data
forest |>
  diagnose() |>             #<2>
  formattable()
```

1.  The `formattable` package is used to create formatted tables and data frames for better data visualization. This package is particularly useful for creating visually appealing and informative tables for data analysis and reporting.

2.  The `diagnose()` is used for data profiling, data quality assessment, and diagnostics.

*\<description\>*

### Forest Area Data

```{r forest_area_basic_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
forest_area |>
  head() |>                    
  formattable()

# The properties of the data
forest_area |>
  diagnose() |>
  formattable()
```

*\<description\>*

### Brazil Loss Data

```{r brazil_loss_basic_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
brazil_loss |>
  head() |>
  formattable()

# The properties of the data
brazil_loss |>
  diagnose() |>
  formattable()
```

*\<description\>*

### Soyabean Usage Data

```{r soyabean_use_basic_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
soyabean_use |>
  head() |>
  formattable()

# The properties of the data
soyabean_use |>
  diagnose() |>
  formattable()
```

*\<description\>*

### Vegetable oil Data

```{r vegetable_oil_basic_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
vegetable_oil |>
  head() |>
  formattable()

# The properties of the data
vegetable_oil |>
  diagnose() |>
  formattable()
```

*\<description\>*

### Deforestation by Source

```{r deforestation_by_source_info, message=FALSE, warning=FALSE}
# Get basic information about the dataset
deforestation_by_source |>
  head() |>
  formattable()

# The properties of the data
deforestation_by_source |>
  diagnose() |>
  formattable()
```

*\<description\>*

### Why we chose this dataset?

The selection of datasets for this project is driven by both technical and analytical considerations, particularly related to data visualization. Moreover, deforestation is a serious environmental problem with far-reaching effects. Understanding its trends, drivers, and impacts is critical for informed decision-making and environmental protection. The dataset from `Our World in Data` provides a reliable data with less junk data and also it contains comprehensive set of variables related to deforestation in this scenario. In the context of data visualization, the technical characteristics of a dataset (cleanliness, structure, consistency) play a key role in creating clear and insightful visualizations.

The dataset is well-structured, making them suitable for editing and visualizing data. It contains various tables like `forest`, `brazil_loss`, `soybean_use`, `forest_area`, `vegetable_oil` and `deforestation_by_source` that allows analysis to target specific aspects of deforestation. These dimensions enable multi-layered analysis and visualization, allowing you to investigate different aspects of a problem. Different variables and data dimensions allow you to create diverse visualizations to consider different aspects of deforestation.

The richness of the dataset provides many opportunities to create effective data visualizations. From time series graphs to choropleth maps to scatter plots, there are many visualization techniques that can effectively convey complex information. In summary, the selection of this dataset is driven by its technical suitability for analysis and its relevance to the critical issue of deforestation.

## Questions

The two questions you want to answer.

### Question 1

\<question here\>

### Question 2

\<question here\>

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

### **Approach for question 1**

### Approach for question 2
