---
title: "Visualizing the Forest Conversion"
subtitle: "Proposal"
format: html
editor: visual
code-annotations: hover
categories: 
  - Data visualization
  - TidyTuesday
---

### Packages Setup

```{r load_pkgs, message=FALSE, warning=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Installed Packages"

### GETTING THE LIBRARIES
if (!require(pacman))
  install.packages(pacman)


pacman::p_load(tidyverse,
               tidytuesdayR,
               dlookr,
               formattable)
```

## Data-set

```{r load_dataset, message=FALSE, warning=FALSE, output=FALSE}
# Getting the Data using the tidytuesdayR package 
deforestation_data <- tidytuesdayR::tt_load(2021, week = 15)

# Getting all the underlying data in the dataset
forest        <- deforestation_data$forest
forest_area   <- deforestation_data$forest_area
brazil_loss   <- deforestation_data$brazil_loss
soybean_use   <- deforestation_data$soybean_use
vegetable_oil <- deforestation_data$vegetable_oil
deforestation_by_source <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/deforestation_by_source.csv')  #<1>
```

1.  We are reading `deforestation_by_source` form raw csv file which is in github since it was not being downloaded from the `tidytuesdayR` package.

The Global Deforestation data-set were published by Hannah Ritchie and Max Roser (2021) in the Our World in Data journal. This data-set contains comprehensive information on global forest cover, deforestation, and related factors. It includes multiple datas:

-   `forest` : Data on net forest conversion and changes in forest cover by country over time.
-   `forest_area` : Information on the change in global forest area as a percent of the global forest area.
-   `brazil_loss` : Details on the loss of Brazilian forest due to specific causes.
-   `soybean_use` : Data on soybean production and use by year and country.
-   `vegetable_oil` : Vegetable oil production by crop type and year.

### Forest Data

```{r forest_basic_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Forest Data Diagnosis Code"

# Get basic information about the dataset
# The properties of the data
forest |>
  diagnose() |>
  formattable()
```

The Forest dataset contains information on the change in forest area every 5 years. With 475 observations, the dataset consists of the 4 variables: entity(country), code, year, and the net forest conversion in hectares. The column `code` has some missing values around 1% and we won't be using that column anyway since we have entity column.

### Forest Area Data

```{r forest_area_basic_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Forest Area Data Diagnosis Code"

# Get basic information about the dataset
# The properties of the data   
forest_area |>
  diagnose() |>
  formattable()
```

The Forest area dataset looks at the change in global forest area as a percent of global forest area amongst a sample of 7846 observations. The data collected consists of 4 variables: entity, code, year, and percent of global forest area. The column `code` has some missing values around 13% and we won't be using that column anyway since we have entity column.

### Brazil Loss Data

```{r brazil_loss_basic_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Brazil Loss Data Diagnosis Code"

# Get basic information about the dataset
# The properties of the data
brazil_loss |>
  diagnose() |>
  formattable()
```

The Brazil loss dataset compares loss of Brazilian forest across different types of forest disturbances.13 observations are analyzed amongst 14 variables which include entity, code, year, commercial crops, flooding, natural disturbances, pasture for livestock, logging for lumber, fire loss, mining, infrastructure, roads, tree plantations, and small scale clearing. And we found no missing values in the data after running `diagnosis()` on top of it.

### Soybean Usage Data

```{r soybean_use_basic_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Soybean Usage Data Diagnosis Code"

# Get basic information about the dataset
# The properties of the data
soybean_use |>
  diagnose() |>
  formattable()
```

The data-set called Soybean usage data-set consists of information relating to soybean consumption and use by year and country. 9897 observations were analyzed across 6 variables: `entity`, `code`, `year`, use for `human food` (e.g., tempeh, tofu), used for `animal food`, and `processed` into vegetable oil/bio-fuel/processed animal feed. The columns `animal_feed` and `processed` are having significant missing values.

### Vegetable oil Data

```{r vegetable_oil_basic_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Vegetable Oil Data Diagnosis Code"

# Get basic information about the dataset
# The properties of the data
vegetable_oil |>
  diagnose() |>
  formattable()
```

In the vegetable oil data-set, 143,832 observations were used to analyze the vegetable oil production by crop type and year. Additionally, the variables consisted of `entity`, `code`, `year`, `crop oil` and `production` which contains the production vegetable oil, and oil production in tons. The column `production` is having significant missing values.

### Deforestation by Source

```{r deforestation_by_source_info, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Deforestation Source Data Diagnosis Code"

# The properties of the data
deforestation_by_source |>
  diagnose() |>
  formattable()
```

10 observations within the deforestation by source dataset compared different farming entities to forest loss by year. The dataset includes four variables: `entity`, `country`, `year`, and `forest loss`. The column `code` has some missing values around 100% and we won't be using that column anyway since we have entity column.

### Why we chose this data-set?

The selection of data-sets for this project is driven by both technical and analytical considerations, particularly related to data visualization. Moreover, deforestation is a serious environmental problem with far-reaching effects. Understanding its trends, drivers, and impacts is critical for informed decision-making and environmental protection. The data-set from `Our World in Data` provides a reliable data with less junk data and also it contains comprehensive set of variables related to deforestation in this scenario.

The data-set is well-structured, making them suitable for editing and visualizing data. It contains various tables like `forest`, `brazil_loss`, `soybean_use`, `forest_area`, `vegetable_oil` and `deforestation_by_source` that allows analysis to target specific aspects of deforestation. Different variables and data dimensions allow us to create diverse visualizations to consider different aspects of deforestation and investigate them.

The richness of the data-set provides many opportunities to create effective data visualizations. From time series graphs to choropleth maps to scatter plots, there are many visualization techniques that can effectively convey complex information. In summary, the selection of this data-set is driven by its technical suitability for analysis and its relevance to the critical issue of deforestation.

## Questions

The two questions to be answered are:

**Question 1:** What does the global forest area look like over past 3 decades, highlighting the trends of forest area conversion?

**Question 2:** How has the production of Soybean in Africa changed over time, and how does it impact the afforestation or deforestation rates?

## Analysis plan

The following are the approaches we will be using for each question.

### Approach for question 1

For this plot, we wanted to visualize how the global area under forests has changed over the years. We wanted to represent this data in a graphical manner and show how reducing forest cover over the years could be related to other ecological changes taking place. To visualize the `TidyTuesday` data-set which was sourced from `Our World in Data`, we are planning to clean it further to only get certain parts that are required using `dplyr`. Out of the data involved we will be mostly focusing in the `forest` data which contains our columns of focus which is `net_forest_conversion`, `year` and `entity`. From our initial analysis, we would not need to create any new variable using the data-set.

And coming the actual visualization part the comparison of the plot will be done through years `1990-2015` as we have slightly better recorded data for these years and using this we should be able to show an overview of changes in forest cover over the past years using `ggplot`. We are planning to create a map plot for the data of forest area and for specific years across the world using the `rnaturalearth`, `ggmap`, `cowplot` and `plotly` packages. The mapping of data from data-set to the map plot can also be done using `geocode()` function which we will be exploring as well.

### Approach for question 2

For this plot, we want to visualize the trend in the production of soybean over the years in the continent of Africa. Three different soybean consumptions are added together to get the overall production. To perform this task, we are going to use the `TidyTuesday` data-set which was sourced from `Our World in Data` and perform data manipulation to obtain a new column showing the total production of soybean. Out of the data involved we will be mostly focusing in the `soyabean_use` data and columns present in the table are - `human_food` ,`animal_feed` and `processed`. We will then use `ggplot` to create a time series plot(line and point graph) to visualize the rate of change in production using the `geom_line()` and `geom_point()` method. Furthermore, we will compare and correlate the increase or decrease in soybean production with the rates of deforestation and afforestation to determine the general trend in both. This data comparison will be done for the years `1990-2013`, as the data-set provides an abundance of useful insights. We also might use other data present in the data-set to get some correlation.

**Variables of focus for both questions:**

| Variable              | Description                                                  | Source Data-set                 |
|------------------|----------------------------------|--------------------|
| entity                | Country                                                      | forest, forest_area and soybean |
| code                  | Country Code                                                 | forest, forest_area and soybean |
| year                  | Year                                                         | forest, forest_area and soybean |
| net_forest_conversion | Net forest conversion in hectares                            | forest                          |
| forest_area           | Percent of global forest area                                | forest_area                     |
| human_food            | Use for human food (tempeh, tofu, etc)                       | soybean                         |
| animal_feed           | Used for animal food                                         | soybean                         |
| processed             | Processed into vegetable oil, biofuel, processed animal feed | soybean                         |

::: callout-note
## Note:

And since these are the approaches we intend to explore and solve the problem statement which we came up with, bits and pieces of the approach might change.
:::
